Estimating duration in the presence of misclassification

### 2024-01-03 ATACCC.tex
Show multiple parameter values in fig

Several features of the \textcite{hakkiOnset} model can be motivated by the desire to apply Stan, a modern and convenient MCMC implementation of the NUTS algorithm (see \cref{E-inc-prev:sec:MCMC}).
The first of these is using a continuous viral load trajectory rather than the more common piecewise linear model.
As NUTS uses gradient information, it is more efficient when the model is differentiable.
This leads to the following model for the viral load at time $t$, $v_i(t)$, in individual $i$ is shown in \cref{ATACCC:fig:viral-load-model} and given by:
\begin{align}
v_i(\tau) = v_{\max,i} \frac{a_i+b_i}{b_ie^{-a_i(\tau-\tau_{\max,i})} + a_ie^{b_i(\tau-\tau_{max,i})}}. \label{ATACCC:eq:viral-load}
\end{align}
Time, $\tau$, is continuous and defined uniquely for each individual such that $\tau = 0$ at the time that individual $i$ was first tested.
$v_{\max,i}$ is the maximum viral load, which occurs at time $\tau_{\max,i}$.
As $\lvert \tau - \tau_{\max} \rvert$ becomes large, the viral load trajectory tends towards exponential growth (before the peak) or decline (after the peak) in viral load at rates $a_i$ and $b_i$ respectively.
% Before the peak, this is because $b(\tau-\tau_{\max,i})$ is small and hence $a\exp(b(\tau-\tau_{\max,i}))$ is negligible.
% After the peak, this is because $b\exp(-a(\tau-\tau_{\max,i}))$ is negligible.
This happens quickly for realistic values of the parameters.
Therefore, the log viral load is approximately piecewise linear (see \cref{ATACCC:fig:viral-load-model} for a representative viral load trajectory).
$\vec{\phi}_i = (v_{\max,i}, a_i, b_i)^T$ are the random effects for individual $i$.
$\tau_{\max,i}$ is inferred independently for each individual because the time from first test to peak is not a biological concept shared between individuals.
\begin{figure}
  \centering \includegraphics{ATACCC/typical_trajectory}
  \caption[Example viral load model trajectory.]{Example viral load trajectory produced by \textcite{hakkiOnset}'s viral load model, using realistic parameter values. The increase in log viral load is approximately linear with a gradient of $a_i$, here 5. The decrease is also linear, with gradient $-b_i$, here -1.4. The peak is at time $\tau_{\max,i}$, here 0, and log viral load $\log v_{\max,i}$, here = 15. The smoothing (lack of linearity) near the peak is visible. \label{ATACCC:fig:viral-load-model}}
  \todo[inline]{Add a few lines to example plot showing the effect of different parameters, move parameters to key}
\end{figure}

>>>>

Several features of the \textcite{hakkiOnset} model can be motivated by the desire to apply Stan, a modern and convenient MCMC implementation of the NUTS algorithm (see \cref{E-inc-prev:sec:MCMC}).
The first of these is using a continuous viral load trajectory rather than the more common piecewise linear model.
As NUTS uses gradient information, it is more efficient when the model is differentiable.
This leads to the following model for the viral load at time $t$, $v_i(t)$, in individual $i$ is shown in \cref{ATACCC:fig:viral-load-model} and given by:
\begin{align}
v_i(\tau) = v_{\max,i} \frac{a_i+b_i}{b_ie^{-a_i(\tau-\tau_{\max,i})} + a_ie^{b_i(\tau-\tau_{max,i})}}. \label{ATACCC:eq:viral-load}
\end{align}
Time, $\tau$, is continuous and defined uniquely for each individual such that $\tau = 0$ at the time that individual $i$ was first tested.
$v_{\max,i}$ is the maximum viral load, which occurs at time $\tau_{\max,i}$.
As $\lvert \tau - \tau_{\max} \rvert$ becomes large, the viral load trajectory tends towards exponential growth (before the peak) or decline (after the peak) in viral load at rates $a_i$ and $b_i$ respectively.
% Before the peak, this is because $b(\tau-\tau_{\max,i})$ is small and hence $a\exp(b(\tau-\tau_{\max,i}))$ is negligible.
% After the peak, this is because $b\exp(-a(\tau-\tau_{\max,i}))$ is negligible.
This happens quickly for realistic values of the parameters.
Therefore, the log viral load is approximately piecewise linear (see \cref{ATACCC:fig:viral-load-model} for a representative viral load trajectory).
$\vec{\phi}_i = (v_{\max,i}, a_i, b_i)^T$ are the random effects for individual $i$.
$\tau_{\max,i}$ is inferred independently for each individual because the time from first test to peak is not a biological concept shared between individuals.
\begin{figure}[tbhp]
  \centering \includegraphics{ATACCC/typical_trajectory}
  \caption[Example viral load model trajectory.]{%
    Example viral load trajectory produced by \textcite{hakkiOnset}'s viral load model, using realistic parameter values.
    The increase and decrease in log viral load are approximately linear with a gradient of $a_i$ and $b_i$ respectively.
    The peak is at time $\tau_{\max,i} = 0$ and log viral load $\log v_{\max,i}$.
    The subscripts are omitted in the key for clarity.
    The smoothing (lack of linearity) near the peak is visible. \label{ATACCC:fig:viral-load-model}
  }
\end{figure}
